Covariate Analysis of ROSMAP Reprocessed counts Regressing out Diagnosis as well for TWAS

### 2019-09-18 code/ROSMAP_RegressDiagnosis_4_TWAS.Rmd
AWS CLI

---
title: Covariate and differential expression analysis of ROSMAP reprocessed counts
  Using Diagnosis as a covariate for TWAS
author: 'JKG - Adapted from: Thanneer Perumal'
output: pdf_document
editor_options:
  chunk_output_type: console
---
Date of analysis update: "`r date()`"

@@

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering, cache=TRUE}
primaryVariable <- c("cogdx", "Diagnosis", "APOE4")
FactorCovariates <- c("Batch", "Sex", "race", "spanish", "cogdx", "Diagnosis", "APOE4")
ContCovariates <- c("RINcontinuous", "RINcontinuous2", "age_death", "pmi", "educ", 
                    "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES",
                    "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", 
                    "PCT_RIBOSOMAL_BASES")
# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES <- data.frame(lapply(COVARIATES,function(x){x <- sapply(x,function(y){str_replace_all(as.character(y),'\\+','')})}))
rownames(COVARIATES) <- METADATA$Sampleid
# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```

Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=9, fig.height=6}
COVARIATES.CORRELATION = CovariateAnalysis::getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left', padding  = unit(c(18,2,2,18), 'mm'))
```

### Explore metadata
```{r data.explore, fig.width = 10, fig.height = 14}
# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Diagnosis, y = RINcontinuous)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + theme(legend.position = 'top')
# AgeAtDeath
p[[2]] = ggplot(COVARIATES, aes(x = Diagnosis, y = age_death)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeOfDeath') + theme(legend.position = 'top')
# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Diagnosis, y = pmi)) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI') + theme(legend.position = 'top')
# Education
p[[4]] = ggplot(COVARIATES, aes(x = Diagnosis, y = educ)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Education') + theme(legend.position = 'top')
# Intronic bases
p[[5]] = ggplot(COVARIATES, aes(x = Diagnosis, y = PCT_INTRONIC_BASES)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Fraction Intronic Bases') + theme(legend.position = 'top')
# Ribosomal bases
p[[6]] = ggplot(COVARIATES, aes(x = Diagnosis, y = PCT_RIBOSOMAL_BASES)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Fraction Ribosomal Bases') + theme(legend.position = 'top')
multiplot(plotlist = p, cols = 2)
```

@@

### Sample clustering
PCA based clustering of samples
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis', cache=FALSE}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E.no.na, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Batch, shape=cogdx, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p
```

Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("Batch", "Sex", "Diagnosis")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E.no.na)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

@@

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates are used to find significant covariates
```{r preadj.covariates, cache=TRUE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E.no.na, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preadj.covariates.plot, fig.width=12, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Normalisation (iterative design)
Since many covariates are correlated, re-normalising and re-adjusting COUNTS with an iterative design matrix
1. Adding Batch and Sex a priori to variable selection
2. Primary variable of interest Diagnosis is excluded from the pool of available covariates for selection
```{r iterative.norm, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Batch', 'Sex');
# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, c(postAdjustCovars, primaryVariable))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
#_# ADD DIAGNOSIS HERE TO REGRESS FOR TWAS
postAdjustCovars = c(postAdjustCovars, "Diagnosis", names(which.max(covariatesEffects))) %>% unique()
loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 20){
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects'))

@@

### Sanity check
```{r residual.adj, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(expr, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(expr))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
PCA of residual data
```{r decompse.normalise.data2.1, fig.height=6, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(expr, scale.=T, center = T)
# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))
p <- ggplot(plotdata, aes(x=PC1, y=PC2)) 
p <- p + geom_point(aes(color=Batch, shape=Diagnosis, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p
```
Tree based clustering of residual data
```{r decompse.normalise.data2.2, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c('Batch', 'Sex', primaryVariable)])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(expr)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

@@

### SVA Adjustments for eQTL analysis
Conditioned on primary variable (Diagnosis) and identified covariates estimate surrogate variables using SVA package
```{r sva.adjust}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Estimate voom weights for dispersion control
cnts = NEW.COUNTS
cnts[is.na(cnts)] = 0
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            design = DESIGN$design,
            weights = VOOM.WEIGHTS$weights)
# Get (null) design matrix
MODEL0 = DESIGN$design[,-grep(primaryVariable[2],colnames(DESIGN$design))] # Get null model by removing variable of interest
MODEL0 = MODEL0[,linColumnFinder(MODEL0)$indepCols]
MODEL1 = DESIGN$design
MODEL1 = MODEL1[,linColumnFinder(MODEL1)$indepCols]
# Calculate the number of surrogate variables (be method)
# NUM.SV = sva::num.sv(CQN.GENE_EXPRESSION$E, MODEL1, method = 'be', seed = 123456, B = 30)
# Get residuals from differential expression model
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(FIT, CQN.GENE_EXPRESSION$E)
# Compute actual variance of all principal components
expr = RESIDUAL.SVA.GENE_EXPRESSION
expr[is.na(expr)] = 0
tmp = svd(expr)
actual.var = tmp$d^2/sum(tmp$d^2)
# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind) %dopar% {
  tmp.residual = t(apply(expr, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])
# Plot variance components
var.comp = data.frame(component = 1:length(actual.var), 
                      residual = actual.var, 
                      permuted.residual = permuted.var) %>%
  tidyr::gather(data, value, -component)
p = ggplot(var.comp %>% filter(component <= round(NUM.SV*1.3)), aes(x = component, y = value, color = data)) + geom_point()
p = p + geom_hline(yintercept=permuted.var[1], linetype = 'dashed') + xlab('Singular Dimension Index')
p = p + geom_vline(xintercept=NUM.SV, linetype = 'dashed') + ylab('Fraction of Variance Explained')
p = p + theme(legend.position = c(0.8, 0.8), legend.title = element_blank())
p
# Estimate surrogate variables
SURR.VAR = sva::sva(CQN.GENE_EXPRESSION$E.no.na, 
                    MODEL1, MODEL0, n.sv = NUM.SV, B = 30)$sv
SURR.VAR = data.frame(SURR.VAR)
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)
# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(cnts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F)
# Fit linear model using new weights and new design
VOOM.SVA.FIT = lmFit(CQN.GENE_EXPRESSION$E, 
                     design = cbind(MODEL1, SURR.VAR),
                     weights = VOOM.SVA.WEIGHTS$weights)
# Residuals after normalisation
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(VOOM.SVA.FIT,
                                                  CQN.GENE_EXPRESSION$E)
# Add variable of interest back to the residuals
varsToAddIn = grep("Diagnosis", colnames(MODEL1), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
```

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective counts with GRCh37 with CQN normalisation (DLPFC)';

@@

stopCluster(cl)
```

@@

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')
synLogin()
knit2synapse::knitfile2synapse(file = "~/TWAS/code/ROSMAP_RegressDiagnosis_4_TWAS.Rmd", 
                               owner = "syn20800296")
```

>>>>

---
title: Covariate Analysis of ROSMAP Reprocessed counts Regressing out Diagnosis as well for TWAS

author: 'JKG - Adapted from: Thanneer Perumal'
output: html_notebook
editor_options:
  chunk_output_type: console
---
Date of analysis update: "`r date()`"

@@

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering, cache=FALSE}
primaryVariable <- c("cogdx", "Diagnosis", "APOE4")
FactorCovariates <- c("Batch", "Sex", "race", "spanish", "cogdx", "Diagnosis", "APOE4")
ContCovariates <- c("RINcontinuous", "RINcontinuous2", "age_death", "pmi", "educ", 
                    "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES",
                    "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", 
                    "PCT_RIBOSOMAL_BASES")
# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES <- data.frame(lapply(COVARIATES,function(x){x <- sapply(x,function(y){str_replace_all(as.character(y),'\\+','')})}))
rownames(COVARIATES) <- METADATA$Sampleid
# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```

Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, cache=FALSE, fig.width=9, fig.height=6}
COVARIATES.CORRELATION = CovariateAnalysis::getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left', padding  = unit(c(18,2,2,18), 'mm')) 
```

### Explore metadata
```{r data.explore, cache=FALSE, fig.width = 10, fig.height = 14}
# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Diagnosis, y = RINcontinuous)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + theme(legend.position = 'top')
# AgeAtDeath
p[[2]] = ggplot(COVARIATES, aes(x = Diagnosis, y = age_death)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeOfDeath') + theme(legend.position = 'top')
# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Diagnosis, y = pmi)) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI') + theme(legend.position = 'top')
# Education
p[[4]] = ggplot(COVARIATES, aes(x = Diagnosis, y = educ)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Education') + theme(legend.position = 'top')
# Intronic bases
p[[5]] = ggplot(COVARIATES, aes(x = Diagnosis, y = PCT_INTRONIC_BASES)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Fraction Intronic Bases') + theme(legend.position = 'top')
# Ribosomal bases
p[[6]] = ggplot(COVARIATES, aes(x = Diagnosis, y = PCT_RIBOSOMAL_BASES)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Fraction Ribosomal Bases') + theme(legend.position = 'top')
multiplot(plotlist = p, cols = 2)
```

@@

### Sample clustering
PCA based clustering of samples
```{r decompse.normalise.data, cache=FALSE, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E.no.na, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Batch, shape=cogdx, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p
```

Tree based clustering of samples
```{r decompse.normalise.data.1, cache=FALSE, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("Batch", "Sex", "Diagnosis")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E.no.na)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

@@

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates are used to find significant covariates
```{r preadj.covariates, cache=FALSE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E.no.na, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preadj.covariates.plot, cache=FALSE, fig.width=12, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Normalisation (iterative design)
Since many covariates are correlated, re-normalising and re-adjusting COUNTS with an iterative design matrix
1. Adding Batch and Sex a priori to variable selection
2. Primary variable of interest Diagnosis is excluded from the pool of available covariates for selection
```{r iterative.norm, cache=FALSE, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Batch', 'Sex');
# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, c(postAdjustCovars, primaryVariable))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
#_# ADD DIAGNOSIS HERE TO REGRESS FOR TWAS
postAdjustCovars = c(postAdjustCovars, "Diagnosis", names(which.max(covariatesEffects))) %>% unique()
loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 20){
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects'))

@@

### Sanity check
```{r residual.adj, cache=FALSE, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(expr, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(expr))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
PCA of residual data
```{r decompse.normalise.data2.1, cache=FALSE, fig.height=6, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(expr, scale.=T, center = T)
# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))
p <- ggplot(plotdata, aes(x=PC1, y=PC2)) 
p <- p + geom_point(aes(color=Batch, shape=Diagnosis, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p
```
Tree based clustering of residual data
```{r decompse.normalise.data2.2, cache=FALSE, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c('Batch', 'Sex', primaryVariable)])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(expr)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

@@

### SVA Adjustments for eQTL analysis
Conditioned on primary variable (Diagnosis) and identified covariates estimate surrogate variables using SVA package
```{r sva.adjust, cache=FALSE }
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Estimate voom weights for dispersion control
cnts = NEW.COUNTS
cnts[is.na(cnts)] = 0
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            design = DESIGN$design,
            weights = VOOM.WEIGHTS$weights)
# Get (null) design matrix
MODEL0 = DESIGN$design[,-grep(primaryVariable[2],colnames(DESIGN$design))] # Get null model by removing variable of interest
MODEL0 = MODEL0[,linColumnFinder(MODEL0)$indepCols]
MODEL1 = DESIGN$design
MODEL1 = MODEL1[,linColumnFinder(MODEL1)$indepCols]
# Calculate the number of surrogate variables (be method)
# NUM.SV = sva::num.sv(CQN.GENE_EXPRESSION$E, MODEL1, method = 'be', seed = 123456, B = 30)
# Get residuals from differential expression model
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(FIT, CQN.GENE_EXPRESSION$E)
# Compute actual variance of all principal components
expr = RESIDUAL.SVA.GENE_EXPRESSION
expr[is.na(expr)] = 0
tmp = svd(expr)
actual.var = tmp$d^2/sum(tmp$d^2)
# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind) %dopar% {
  tmp.residual = t(apply(expr, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])
# Plot variance components
var.comp = data.frame(component = 1:length(actual.var), 
                      residual = actual.var, 
                      permuted.residual = permuted.var) %>%
  tidyr::gather(data, value, -component)
p = ggplot(var.comp %>% filter(component <= round(NUM.SV*1.3)), aes(x = component, y = value, color = data)) + geom_point()
p = p + geom_hline(yintercept=permuted.var[1], linetype = 'dashed') + xlab('Singular Dimension Index')
p = p + geom_vline(xintercept=NUM.SV, linetype = 'dashed') + ylab('Fraction of Variance Explained')
p = p + theme(legend.position = c(0.8, 0.8), legend.title = element_blank())
p
# Estimate surrogate variables
SURR.VAR = sva::sva(CQN.GENE_EXPRESSION$E.no.na, 
                    MODEL1, MODEL0, n.sv = NUM.SV, B = 30)$sv
SURR.VAR = data.frame(SURR.VAR)
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)
# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(cnts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F)
# Fit linear model using new weights and new design
VOOM.SVA.FIT = lmFit(CQN.GENE_EXPRESSION$E, 
                     design = cbind(MODEL1, SURR.VAR),
                     weights = VOOM.SVA.WEIGHTS$weights)
# Residuals after normalisation
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(VOOM.SVA.FIT,
                                                  CQN.GENE_EXPRESSION$E)
# Add variable of interest back to the residuals
varsToAddIn = grep("Diagnosis", colnames(MODEL1), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
```

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=TRUE}
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective counts with GRCh37 with CQN normalisation (DLPFC)';

@@

#stopCluster(cl)
```

@@

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
library(synapser)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "/home/jgockley/TWAS/code/ROSMAP_RegressDiagnosis_4_TWAS.Rmd",
                                          parentId ="syn20798011",
                                          folderName = 'ROSMAP')
```
